---
output: html_document
---

The following examples illustrates the reading of the meassured tree locations.

### Workflow
The workflow is straight forward. Read the file and adjust the content.


### Reading the data
Reading this data format is straight forward.
```{r}
wd <- "D:/active/moc/pt/examples/data/topo_survey/"

df <- read.table(paste0(wd, "topo_survey.csv"), 
                 header = TRUE, sep = ";", dec = ".")
head(df)
```


### Adjusting the content
While the angle and distance is already stored in individual columns, the from/to tree information is in a single column which is not flexible. Hence, this information will be split into two and the column names will be adjusted.
```{r}
# Adjust column names
colnames(df) <- c("Original", "Distance", "Angle")
head(df)

# Split first column into two new ones; keep the original single column version
# just for cross-checking purposes.
split_pos <- regexpr("->",  df$Original)
df$From <- substr(as.character(df$Original), 1, split_pos-2)
df$To <- substr(as.character(df$Original), split_pos+2, 
                nchar(as.character(df$Original)))

head(df)
```


