---
output: html_document
---

The following examples illustrates adds a variation to the [From Distance to Coordinates](http://moc.environmentalinformatics-marburg.de/doku.php?id=courses:bsc:project-thesis-geoei:lecture-notes:gis:pt-ge-ln-gis-01-2000) workflow by not using a fixed base point but meassuring from one point to the next one (i.e. as a path way).

### Workflow
The workflow is quite the same as in [From distance to coordinates](http://moc.environmentalinformatics-marburg.de/doku.php?id=courses:bsc:project-thesis-geoei:lecture-notes:gis:pt-ge-ln-gis-01-2000) and we use exactely the same helper functions. The exception is, that we use the just computed map coordinates as the base point for the next computation.

In the following, we will not show file reading or the helper functions, again, but just show one example of a loop. The data frame used for this example is the one from [Reading tree observation sheets](http://moc.environmentalinformatics-marburg.de/doku.php?id=courses:bsc:project-thesis-geoei:lecture-notes:bdh:pt-ge-ln-bdh-01-2060_trees).

```{r, echo = FALSE, warnings=FALSE}
library(rgdal)
library(raster)
library(sp)

wd <- "D:/active/moc/pt/examples/data/topo_survey/"
df <- read.table(paste0(wd, "topo_survey.csv"), 
                 header = TRUE, sep = ";", dec = ".")

# Adjust column names
colnames(df) <- c("Original", "Distance", "Angle")
head(df)

# Split first column into two new ones; keep the original single column version
# just for cross-checking purposes.
split_pos <- regexpr("->",  df$Original)
df$From <- substr(as.character(df$Original), 1, split_pos-2)
df$To <- substr(as.character(df$Original), split_pos+2, 
                nchar(as.character(df$Original)))

# Function to convert degrees to radians
deg2Rad <- function(degrees) degrees * pi / 180

# Function to convert direction and distance to x and y components in meter
dirDis2XY <- function(dir, dis) {
  u <- -dis * sin(deg2Rad(dir))
  x <- round(-u, 2)
  v <- -dis * cos(deg2Rad(dir))
  y <- round(-v, 2)
  return(as.data.frame(cbind(x, y)))
}

# Function to convert geographic coordinates to UTM-32N and transform the date
# from WGS84 to GRS80
latLong2UTM32 <- function(bpx, bpy){
  p <- data.frame(x = bpx,
                  y = bpy)
  coordinates(p) <- ~x+y
  projection(p) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  p <- spTransform(p, CRS("+proj=utm +zone=32 +ellps=GRS80 +units=m +north"))
  return(coordinates(p))  
}

# Function to calculate the UTM-32N coordinates for your point of interest
# Arg
# - bpx: Longitude of your GPS measured base point
# - bpy: Latitude of your GPS measured base point
# - dir: direction from your base point to your point of interest
# - dis: distance from your base point to your point of interest
calcNewPoint <- function(bpx, bpy, dir, dis){
  offset <- dirDis2XY(dir, dis)
  poi <- data.frame(x = bpx + offset$x,
                  y = bpy + offset$y)
  return(poi)
}
```

### Computing map coordinates from a path way
We use an arbitrary first base point in geographical coordinates for this example. This basepoint is converted to UTM 32 north (GRS80). 

The map coordinates of this base point are stored in a data frame called mapp. Starting with line one in the path way dataset, which gives us the distance and angle to some point N from this initial base point, we compute the coordinates of N and save it in the second line of the mapp. The loop over our path way continoues and the next time, the position of N is used as basepoint and so on.
```{r}
x <- 8.913101 # base point x coordinate in geographical coordinates
y <- 51.15143 # base point y coordinate in geographical coordinates

# convert coordinates to UTM-32N
coordinates_utm <- latLong2UTM32(x,y)
x <- coordinates_utm[1]
y <- coordinates_utm[2]

mapp <- data.frame(PlotID = "TP",
                   X = x,
                   Y = y)
for(i in seq(nrow(df))){
  act_pos <- grep(df$From[i], mapp$PlotID)[1]
  new_point <- calcNewPoint(mapp$X[act_pos], mapp$Y[act_pos], df$Angle[i], df$Distance[i])
  newp <- data.frame(PlotID = df$To[i],
                     X = new_point$x,
                     Y = new_point$y)
  mapp <- rbind(mapp, newp)  
}
head(mapp)
```

You can try the following on your PC to check the locations.
```{r, eval = FALSE}
# You can try this on your PC to check the locations.
library(mapview)
coordinates(mapp) <- ~X+Y
projection(mapp) <- CRS("+proj=utm +zone=32 +ellps=GRS80 +units=m +north")
mapview(mapp)
```


